---
output: github_document
---

```{r, echo = FALSE, results = 'asis'}
suppressPackageStartupMessages(suppressWarnings(library(huxtable)))
print_html(hux_logo())

```

<!-- README.md is generated from README.Rmd. Please edit that file -->
<br>

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  echo = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```


<!-- badges: start -->
[![R-universe Badge](https://hughjonesd.r-universe.dev/badges/huxtable)](https://hughjonesd.r-universe.dev/huxtable)
[![CRAN Status Badge](https://www.r-pkg.org/badges/version/huxtable)](https://cran.r-project.org/package=huxtable)
[![CRAN Downloads Per Month](https://cranlogs.r-pkg.org/badges/huxtable)](https://CRAN.R-project.org/package=huxtable)
[![R build status](https://github.com/hughjonesd/huxtable/workflows/R-CMD-check/badge.svg)](https://github.com/hughjonesd/huxtable/actions)
[![Coverage Status](https://img.shields.io/codecov/c/github/hughjonesd/huxtable/master.svg)](https://app.codecov.io/github/hughjonesd/huxtable?branch=master)
<!-- badges: end -->
  
Huxtable is an R package to create styled tables in multiple output formats, with a friendly, modern interface. Features include:

* Control over text styling, number format, background colour, borders, padding and alignment.
* Table cells can span multiple rows and/or columns.
* Table manipulation via standard R subsetting, or using `dplyr`. 
* Automatic formatting for knitr/rmarkdown documents.
* `huxreg()` function for quick creation of regression tables.
* Output to HTML, LaTeX, RTF, and Microsoft Word/Excel/Powerpoint, using the `officer` 
  and `openxlsx` packages. 
* Quick one-liners to print data frames into a new PDF, HTML page, RTF or Microsoft document.
* Formatted table display in the R console, including borders, colour, and text styles.

# Installing

To install from [r-universe](https://r-universe.dev):

```{r, eval = FALSE}
install.packages("huxtable", repos = c("https://hughjonesd.r-universe.dev", 
				       "https://cloud.r-project.org"))
```

To install from CRAN:

```{r, eval = FALSE}
install.packages('huxtable')
```

To install the latest version from github: 

```{r, eval = FALSE}
install.packages('remotes')
remotes::install_github('hughjonesd/huxtable')
```

# Learning more

Check out [the website](https://hughjonesd.github.io/huxtable/), read the
[documentation](https://hughjonesd.github.io/huxtable/reference/index.html) or 
read the vignette in [HTML](https://hughjonesd.github.io/huxtable/huxtable-html.html) 
or [PDF](https://hughjonesd.github.io/huxtable/huxtable-pdf.pdf).


