---
output: github_document
---

```{r, echo = FALSE, results = 'asis'}
suppressPackageStartupMessages(library(huxtable))
print_html(hux_logo())

```

<!-- README.md is generated from README.Rmd. Please edit that file -->
<br>

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  echo = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```



[![CRAN Status Badge](http://www.r-pkg.org/badges/version/huxtable)](https://cran.r-project.org/package=huxtable)
[![Travis-CI Build Status](https://travis-ci.org/hughjonesd/huxtable.svg?branch=master)](https://travis-ci.org/hughjonesd/huxtable)
[![CRAN Downloads Per Month](http://cranlogs.r-pkg.org/badges/huxtable)](https://CRAN.R-project.org/package=huxtable)
[![Coverage Status](https://img.shields.io/codecov/c/github/hughjonesd/huxtable/master.svg)](https://codecov.io/github/hughjonesd/huxtable?branch=master)

Huxtable is an R package for creating HTML and LaTeX tables. It provides similar 
functionality to xtable, but does more, with a simpler interface. Huxtable includes
a `huxreg()` function for quick creation of regression tables, and allows
output to Microsoft Word or Powerpoint using the [`officer` package](https://cran.r-project.org/package=officer).

# Installing

To install from CRAN:

```{r, eval = FALSE}
install.packages('huxtable')
```

To install from github: 

```{r, eval = FALSE}
install.packages('devtools')
devtools::install_github('hughjonesd/huxtable')
```

# Learning more

Check out [the website](https://hughjonesd.github.io/huxtable), read the
[documentation](https://hughjonesd.github.io/huxtable/reference/index.html) or 
read the vignette in [HTML](https://hughjonesd.github.io/huxtable/huxtable.html) 
or [PDF](https://hughjonesd.github.io/huxtable/huxtable.pdf).
