---
title: "Quick guide to functionality"
author: "David Hugh-Jones"
date: "20/02/2017"
output:
  html_document: default
  pdf_document:
    includes:
      in_header: placeins-header.tex
    keep_tex: yes
---


```{r setup, echo = FALSE}
knitr::asis_output('', meta = rmarkdown::latex_dependency('placeins'))
library(huxtable)
is_latex <- knitr::opts_knit$get('rmarkdown.pandoc.to') == 'latex'

fbr <- function() if (is_latex) knitr::asis_output('\\FloatBarrier')
knitr::knit_hooks$set(
  barrier = function(before, options, envir) {
    if (! before) fbr()
  }
)
knitr::opts_chunk$set(echo = TRUE, barrier = TRUE)

```

# Basic usage

## Creating a huxtable

```{r}
ht_orig <- huxtable(a = c('Parsley', 'Sage', 'Rosemary', 'Thyme'), b = 10 ^ (2:5)) 
width(ht_orig) <- 0.25
ht_orig <- set_all_borders(ht_orig, 1:nrow(ht_orig), 1:ncol(ht_orig), 1)
ht_orig
```

## Table position

```{r}
ht <- ht_orig
position(ht) <- 'left'
ht
```

## Cell alignment

```{r}
ht <- ht_orig
align(ht)[,1] <- 'left'
ht

```

## Vertical alignment

```{r}
ht <- ht_orig
valign(ht)[,1] <- 'top'
ht

```

## Column span

```{r}

ht <- ht_orig
background_color(ht)[1,1] <- 'green'
colspan(ht)[1,1] <- 2
ht   

```

## Row span

```{r}
ht <- ht_orig
ht[1,1] <- 'A rather long block of text'
background_color(ht)[1,1] <- 'green'
rowspan(ht)[1,1] <- 2
ht   

```


## Row and column span

```{r}
ht <- ht_orig
ht[1,1] <- 'A rather long block of text'
background_color(ht)[1,1] <- 'green'
rowspan(ht)[1,1] <- 2
colspan(ht)[1,1] <- 2
ht   

```

## Table width

```{r}
ht <- ht_orig
width(ht) <- 0.5
ht

```

## Column width

```{r}
ht <- ht_orig
col_width(ht) <- c(.8, .2)
ht

```

## Row height

```{r}
ht <- ht_orig
height(ht) <- 0.2; 
if (! is_latex) height(ht) <- '100px' # need a specific height for row heights to work in HTML
row_height(ht) <- c(.4, .2, .2, .2) * 0.2
ht

```

## Borders

```{r}
ht <- ht_orig
ht <- set_all_borders(ht, 1:4, 1:2, 0)
top_border(ht) <- c(2, 0, 0, 1)
bottom_border(ht)[4,] <- 2 
ht

```

## Text color, bold, italic, font size

```{r}
ht <- ht_orig
text_color(ht)[,2] <- 'red'
bold(ht)[1,] <- TRUE
italic(ht)[3,] <- TRUE
font_size(ht)[2,] <- 14
font(ht)[4,] <- if (is_latex) 'phv' else 'times' 
ht

```

## Replace NAs

```{r}
ht <- ht_orig
ht[2,] <- NA
ht
na_string(ht) <- '--'
ht

```


# Joining, subsetting and manipulating huxtables

## Subsets

```{r}
ht <- ht_orig
bottom_border(ht)[c(1,4),] <- 1
background_color(ht)[1,] <- 'wheat'

ht[1:3,1]
```

## Joining

```{r}

ht2 <- ht_orig
italic(ht2) <- TRUE
rbind(ht, ht2)
cbind(ht, ht2)
rbind(c("Oregano", 300), ht)
```

## Transpose

```{r}
colspan(ht)[1, 1] <- 2
ht
t(ht)
```


# Advanced usage


## Number formatting

```{r}
ht <- ht_orig
ht[,2] <- ht[,2] + rnorm(4)
number_format(ht)[2,] <- 2
number_format(ht)[3,] <- '%011.4f'
number_format(ht)[4,] <- list(function(x) prettyNum(round(x, 3), big.mark = ','))

ht

```

## Cell rotation

```{r}
ht <- ht_orig
height(ht) <- 0.2 # necessary
if (! is_latex) height(ht) <- '300px'
col_width(ht) <- c(.25, .75)
rotation(ht)[,1] <- 90
ht

```

## Caption

```{r}
ht <- ht_orig
caption(ht) <- 'A simple table'
ht

```

## Caption below

```{r}
ht <- ht_orig
caption(ht) <- 'A simple table'
caption_pos(ht) <- 'bottom'
ht

```

## Label (LaTeX only)

```{r}
ht <- ht_orig
caption(ht) <- 'Captions are required for labels to work'
label(ht) <- 'tab:mytable'
knitr::asis_output('If this is LaTeX we can see a reference to table \\ref{tab:mytable}.')
```

NB: references may not work if knitting PDFs in Rstudio. 

