---
title: "Quick guide to functionality"
author: "David Hugh-Jones"
date: "2/03/2017"
output:
  pdf_document:
    includes:
      in_header: placeins-header.tex
    keep_tex: yes
  html_document: null
---


```{r setup, echo = TRUE}
knitr::asis_output('', meta = rmarkdown::latex_dependency('placeins'))
library(huxtable)
is_latex <- knitr::opts_knit$get('rmarkdown.pandoc.to') == 'latex'

knitr::knit_hooks$set(
  barrier = function(before, options, envir) {
    if (! before && is_latex) knitr::asis_output('\\FloatBarrier')
  }
)
knitr::opts_chunk$set(echo = TRUE, barrier = TRUE)

```

# Basic usage

## Creating a huxtable

```{r}
ht_orig <- huxtable(a = c('Parsley', 'Sage', 'Rosemary', 'Thyme'), b = 10 ^ (2:5)) 
width(ht_orig) <- 0.25
ht_orig <- set_all_borders(ht_orig, 1:nrow(ht_orig), 1:ncol(ht_orig), 1)
ht_orig
```

## Table position

```{r}
ht <- ht_orig
position(ht) <- 'left'
ht
```

## Cell alignment

```{r}
ht <- ht_orig
align(ht)[,1] <- 'left'
ht

```

## Vertical alignment

```{r}
ht <- ht_orig
valign(ht)[,1] <- 'top'
ht

```

## Column span

```{r}

ht <- ht_orig
background_color(ht)[1,1] <- 'green'
colspan(ht)[1,1] <- 2
ht   

```

## Row span

```{r}
ht <- ht_orig
ht[1,1] <- 'A rather long block of text'
background_color(ht)[1,1] <- 'green'
rowspan(ht)[1,1] <- 2
ht   

```



## Table width

```{r}
ht <- ht_orig
width(ht) <- 0.5
ht

```

## Column width

```{r}
ht <- ht_orig
col_width(ht) <- c(.8, .2)
ht

```

## Row height

```{r}
ht <- ht_orig
height(ht) <- 0.2; 
if (! is_latex) height(ht) <- '100px' # need a specific height for row heights to work in HTML
row_height(ht) <- c(.4, .2, .2, .2) * 0.2
ht

```

## Borders

```{r}
ht <- ht_orig
ht <- set_all_borders(ht, 1:4, 1:2, 0)
top_border(ht) <- c(2, 0, 0, 1)
bottom_border(ht)[4,] <- 2 
ht

```


## Text color, bold, italic, font size

```{r}
ht <- ht_orig
text_color(ht)[,2] <- 'red'
bold(ht)[1,] <- TRUE
italic(ht)[3,] <- TRUE
font_size(ht)[2,] <- 14
font(ht)[4,] <- if (is_latex) 'phv' else 'times' 
ht

```


## Number formatting

```{r}
ht <- ht_orig
ht[,2] <- ht[,2] + rnorm(4)
number_format(ht)[2,] <- 2
number_format(ht)[3,] <- '%011.4f'
number_format(ht)[4,] <- list(function(x) prettyNum(round(x, 3), big.mark = ','))

ht

```


## Replace NAs

```{r}
ht <- ht_orig
ht[2,] <- NA
ht
na_string(ht) <- '--'
ht

```


## Quick themes

```{r}
ht <- ht_orig
ht <- rbind(c("Herb", "Amount"), ht)
theme_minimal(ht)
```

```{r}
theme_striped(ht, header_col = FALSE)
```

```{r}
theme_article(ht)
```


# Converting other objects to huxtables

```{r}
defaults <- function(ht) {
  bottom_border(ht)[1,] <- 1
  background_color(ht)[seq(1, nrow(ht), 2), ] <- grey(.9)
  ht
}
```

```{r}
defaults(as_hux(letters[1:5]))
defaults(as_hux(matrix(letters[1:10], 2)))
```

```{r}
data(mtcars)
defaults(as_hux(mtcars[1:10,]))
```

```{r}
car_table <- xtabs(~ cyl + gear, mtcars)
defaults(as_hux(car_table))
```

```{r}
car_ft <- ftable(cyl ~ gear + vs, mtcars)
defaults(as_hux(car_ft))
```


# Joining, subsetting and manipulating huxtables

## Subsets

```{r}
ht <- ht_orig
bottom_border(ht)[c(1,4),] <- 1
background_color(ht)[1,] <- 'wheat'

ht[1:3,1]
```

## Joining

```{r}

ht2 <- ht_orig
italic(ht2) <- TRUE
rbind(ht, ht2)
cbind(ht, ht2)
rbind(c("Oregano", 300), ht)
```

## Transpose

```{r}
colspan(ht)[1, 1] <- 2
ht
t(ht)
```


# Advanced features

## Printing on screen

```{r}
ht <- ht_orig
# Basic data:
print(ht)

# Pretty print:
print_screen(ht)
```


## Padding

```{r}
ht <- ht_orig
ht <- set_all_padding(ht, 1:4, 1:2, 0)
left_padding(ht)[,1] <- 40
right_padding(ht)[,2] <- 40
top_padding(ht) <- 20
bottom_padding(ht) <- 20
ht

```

## Cell rotation

```{r}
ht <- ht_orig
height(ht) <- 0.2 # necessary
if (! is_latex) height(ht) <- '300px'
col_width(ht) <- c(.25, .75)
rotation(ht)[,1] <- 90
ht

```


## Border width

```{r}
ht <- ht_orig
bottom_border(ht)[1,] <- 3
ht

```


## Caption

```{r}
ht <- ht_orig
caption(ht) <- 'A simple table'
ht

```


## Caption below

```{r}
ht <- ht_orig
caption(ht) <- 'A simple table'
caption_pos(ht) <- 'bottom'
ht

```


## Label

```{r}
ht <- ht_orig
caption(ht) <- 'Captions are required for labels to work'
label(ht) <- 'tab:mytable'
ht
knitr::asis_output('If this is LaTeX we can see a reference to table \\ref{tab:mytable}.')
```


NB: references may not work if knitting PDFs in Rstudio. 


# Testing corner cases

## Heights and padding

```{r}
ht <- ht_orig

row_height(ht) <- c(.2, .1, .1, .1)
# top_padding(ht) <- 10
# bottom_padding(ht) <- 20
ht
```

## Row and column span

```{r}
ht <- ht_orig
ht[1,1] <- 'A rather long block of text'
background_color(ht)[1,1] <- 'green'
rowspan(ht)[1,1] <- 2
colspan(ht)[1,1] <- 2
ht   

```

