---
title: "Regression Tables with huxreg"
author: "David Hugh-Jones"
date: '`r Sys.Date()`'
output: 
  html_document: default
  pdf_document:
    includes:
      in_header: placeins-header.tex
vignette: >
  %\VignetteIndexEntry{Regression Tables with huxreg}  
  %\VignetteEngine{knitr::rmarkdown}  
  \usepackage[utf8]{inputenc}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(huxtable)

is_latex <- guess_knitr_output_format() == 'latex'
knitr::knit_hooks$set(
  barrier = function(before, options, envir) {
    if (! before && is_latex) knitr::asis_output('\\FloatBarrier')
  }
)

if (is_latex) knitr::opts_chunk$set(barrier = TRUE)
```

## Regression tables with `huxreg`

From version 0.2, huxtable includes the function `huxreg` to build a table of regressions.

`huxreg` can be called with a list of models. These models can be of any class which has a `tidy` method defined
in the [broom](https://cran.r-project.org/?package=broom) package. The method should return a list of regression
coefficients with names `term`, `estimate`, `std.error` and `p.value`. That covers most standard regression packages.

Let's start by running some regressions to predict a diamond's price.

```{r}
data(diamonds, package = 'ggplot2')

lm1 <- lm(price ~ carat + depth, diamonds)
lm2 <- lm(price ~ depth + factor(color, ordered = FALSE), diamonds)
lm3 <- lm(log(price) ~ carat + depth, diamonds)
```

Now, we call `huxreg` to display the regression output side by side.

```{r}

huxreg(lm1, lm2, lm3)
```

The basic output includes estimates, standard errors and summary statistics. 

Some of those variable names are hard to read. We can change them by specifying a named list of variables in the
`coefs` argument, like this:


```{r}
color_names <- paste0('factor(color, ordered = FALSE)', LETTERS[5:10])
names(color_names) <- paste('Color:', LETTERS[5:10])

huxreg(lm1, lm2, lm3, coefs = c('Carat' = 'carat', 'Depth' = 'depth', color_names))
```

Alternatively, since the output from `huxreg` is just a huxtable, we could just edit its contents directly before we
print it:

```{r}
diamond_regs <- huxreg(lm1, lm2, lm3)
diamond_regs[seq(8, 18, 2), 1] <- paste('Color:', LETTERS[5:10])
diamond_regs
```

Of course, we aren't limited to just changing names. We can also make our table prettier. Let's add the "article" theme,
and a vertical stripe for background colour, tweak a few details like font size, and add a caption. All of these are just 
standard huxtable commands.

```{r}
library(dplyr)
diamond_regs                                                         %>% 
      theme_article                                                  %>% 
      set_background_color(1:nrow(diamond_regs), evens, grey(.95)) %>% 
      set_font_size(final(), 1, 9)                                   %>% 
      set_bold(final(), 1, FALSE)                                    %>%
      set_top_border(final(), 1, 1)                                  %>%
      set_caption('Linear regressions of diamond prices')

```

We could do more, like changing the `number_format` of N to not display decimals. But let's explore what else `huxreg` 
itself can do. 

By default, standard errors are shown below coefficient estimates. To display them in a column to the right, use
`error_pos = 'right'`:

```{r}
huxreg(lm1, lm3, error_pos = 'right')
```


This will give column headings a column span of 2.

To display standard errors in the same cell as estimates, use `error_pos = 'same'`:

```{r}
huxreg(lm1, lm3, error_pos = 'same')
```


You can change the default column headings by giving names to your models:

```{r}
huxreg('Price' = lm1, 'Log price' = lm3)
```


To display a particular row of summary statistics, use the `statistics` parameter. This should be a character vector.
Valid values are anything returned from your models by `broom::glance`. Another valid value is `"nobs"`, which 
returns the number of observations from the regression. If the `statistics` vector has names, these will be used for
row headings:

```{r}
broom::glance(lm1)
huxreg(lm1, lm3, statistics = c('# observations' = 'nobs', 'R squared' = 'r.squared', 'F statistic' = 'statistic',
  'P value' = 'p.value'))
```


By default, `huxreg` displays significance stars. You can alter the symbols used and significance levels 
with the `stars` parameter, or set `stars = NULL` to turn off significance stars completely.

```{r}
huxreg(lm1, lm3, stars = c(`*` = 0.1, `**` = 0.05, `***` = 0.01)) # a little boastful?
huxreg(lm1, lm3, stars = NULL) 
```

You aren't limited to displaying standard errors of the estimates. If you prefer, you can display t statistics or
p values, using the `error_format` option. Any column from `tidy` can be used by putting it in curly brackets:

```{r}
huxreg(lm1, lm3, error_format = '({statistic})')
huxreg(lm1, lm3, error_format = '({p.value})')
```


Or you can display confidence intervals. Use `ci_level` to set the confidence level for the interval, then 
use `{conf.low}` and `{conf.high}` in `error_format`:

```{r}
huxreg(lm1, lm3, ci_level = .99, error_format = '{conf.low} to {conf.high}')
```


To change the footnote, use `note`. If `note` contains the string `"{stars}"` it will be replaced by a description of the
significance stars used. If you don't want a footnote, just set `note = NULL`.

```{r}
huxreg(lm1, lm3, note = 'Linear regressions on diamond price. {stars}.')
```


To change number formatting, set the `number_format` parameter. This works the same as the `number_format` property
for a huxtable - if it is numeric, numbers will be rounded to that many decimal places; if it is character, it will be
taken as a format to the base R `sprintf` function; if it is a function, the function will be called to format the number. 
`huxreg` tries to be smart and to format summary statistics like `nobs` as integers.

```{r}
huxreg(lm1, lm3, number_format = 2)
```


Lastly, if you want to bold all significant coefficients, set the parameter `bold_signif` to a maximum significance level:

```{r}
huxreg(lm1, lm3, bold_signif = 0.05)
```

