---
title: "Introduction to Huxtable"
author: "David Hugh-Jones"
date: '`r Sys.Date()`'
output:
  html_document:
    toc: yes
    toc_depth: 2
  pdf_document:
    includes:
      in_header: placeins-header.tex
    toc: yes
    toc_depth: 2
vignette: >
  %\VignetteIndexEntry{Introduction to Huxtable}   
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}   
---


```{r setup, echo = FALSE, warning = FALSE, message = FALSE}

library(knitr)
library(dplyr)
library(huxtable)
options(huxtable.knit_print_df = FALSE)

is_latex <- guess_knitr_output_format() == 'latex'
# is_latex <- TRUE
knitr::knit_hooks$set(
  barrier = function(before, options, envir) {
    if (! before && is_latex) knitr::asis_output('\\FloatBarrier')
  }
)

if (is_latex) knitr::opts_chunk$set(barrier = TRUE)

```


```{r, echo = FALSE}
huxtable::hux_logo(latex = is_latex, html = ! is_latex)
```

# Introduction

## About this document

This is the introductory vignette for the R package 'huxtable', version `r packageVersion('huxtable')`. A current
version is available on the web in [HTML](https://hughjonesd.github.io/huxtable/huxtable.html) or 
[PDF](https://hughjonesd.github.io/huxtable/huxtable.pdf) format.

## Huxtable

Huxtable is a package for creating *text tables*. It is powerful, but easy to use. It is meant to be a
replacement for packages like xtable, which is useful but not always very user-friendly. Huxtable's features
include:

* Export to LaTeX, HTML, Microsoft Word, Microsoft Excel, Microsoft Powerpoint, RTF and Markdown
* Easy integration with knitr and rmarkdown documents
* Formatted on-screen display
* Multirow and multicolumn cells
* Fine-grained control over cell background, spacing, alignment, size and borders
* Control over text font, style, size, colour, alignment, number format and rotation
* Table manipulation using standard R subsetting, or dplyr functions like `filter` and `select`
* Easy conditional formatting based on table contents
* Quick table themes
* Automatic creation of regression output tables with the `huxreg` function

We will cover all of these features below.

## Installation

If you haven't already installed huxtable, you can do so from the R command line:

```{r, eval = FALSE}
install.packages('huxtable')
```


## Getting started

A huxtable is a way of representing a table of text data in R. You already know that R can represent a
table of data in a data frame. For example, if `mydata` is a data frame, then `mydata[1, 2]` represents the
the data in row 1, column 2, and `mydata$start_time` is all the data in the column called `start_time`.

A huxtable is just a data frame with some extra properties. So, if `myhux` is a huxtable, then `myhux[1, 2]`
represents the data in row 1 column 2, as before. But this cell will also have some other properties - for
example, the font size of the text, or the colour of the cell border.

To create a table with huxtable, use the function `huxtable`, or `hux` for short. This works very much like
`data.frame`.

```{r}
library(huxtable)
ht <- hux(
        Employee     = c('John Smith', 'Jane Doe', 'David Hugh-Jones'), 
        Salary       = c(50000, 50000, 40000),
        add_colnames = TRUE
      )
```


If you already have your data in a data frame, you can convert it to a huxtable with `as_hux`.

```{r}
data(mtcars)
car_ht <- as_hux(mtcars)
```

You can input your data row-by-row using the `tribble_hux` function. This can be more readable:

```{r}
tribble_hux(
  ~Employee,          ~Salary,
  "John Smith",       50000,
  "Jane Doe",         50000,
  "David Hugh-Jones", 40000,
  add_colnames = TRUE
)
```

If you look at a huxtable in R, it will print out a simple representation of the data. Notice that we've added the
column names to the data frame itself, using the `add_colnames` argument to `hux`. We're going to print them out, so they need to be part of the actual table. **NB:** This means that row 1 of your data will be row 2 of the huxtable, and the
column names of your data will be the new row 1.

```{r, results = 'markup'}
print_screen(ht)     # on the R command line, you can just type "ht"
```

To print a huxtable out using LaTeX or HTML, just call `print_latex` or `print_html`. In knitr
documents, like this one, you can simply evaluate the hux. It will know what format to print itself
in.

```{r}
ht
```


# Changing the look and feel

## Huxtable properties

The default output is a very plain table. Let's make it a bit smarter. We'll make the table headings
bold, draw a line under the header row, and add some horizontal space to the cells. We also need to
change that default number formatting to look less scientific.

To do this, we need to set some **properties** on the table cells. You set properties by assigning
to the property name, just as you assign `names(x) <- new_names` in base R. The following commands
assign the value 10 to the `right_padding` and `left_padding` properties, for all cells in `ht`:

```{r}
right_padding(ht) <- 10
left_padding(ht)  <- 10
```

Similarly, we can set the `number_format` property to change how numbers are displayed in 
cells:

```{r}
number_format(ht) <- 2    # 2 decimal places
```

To assign properties to just some cells, you use subsetting, as in base R. So, to make the
first row of the table **bold** and give it a bottom border, we do:

```{r}
bold(ht)[1, ]          <- TRUE
bottom_border(ht)[1, ] <- 1
```

After these changes, our table looks smarter:

```{r}
ht
```


So far, all these properties have been set at cell level. Different cells can have different
alignment, text formatting and so on. By contrast, `caption` is a table-level property. It only
takes one value, which sets a table caption.

```{r}

caption(ht) <- 'Employee table'
ht

```


As well as cell properties and table properties, there is also one row property, `row_height`, and
one column property, `col_width`.

The table below shows a complete list of properties. Most properties work the same for LaTeX and
HTML, though there are some exceptions.

```{r, echo = FALSE}
sides <- c('left_', 'right_', 'top_', 'bottom_')
props <- list()
props[['Cell_Text']] <- sort(c('font', 'text_color', 'wrap', 'bold', 'italic', 'font', 'font_size', 'na_string', 'escape_contents', 'number_format', 'rotation'))

props[['Cell']] <- sort(c('align', 'valign', 'rowspan', 'colspan', 'background_color', 
      paste0(sides, 'border'), paste0(sides, 'border_color'), paste0(sides, 'border_style'), 
      paste0(sides, 'padding')))
props[['Row']]    <- 'row_height'
props[['Column']] <- 'col_width'
props[['Table']]  <- sort(c('width', 'height', 'position', 'caption', 'caption_pos', 'tabular_environment', 'label', 'latex_float'))

maxl <- max(sapply(props, length))
props <- lapply(props, function(x) c(x, rep('', maxl - length(x))))

ss_font <- if (guess_knitr_output_format() == 'latex') 'cmtt' else 'courier'

prop_hux <- hux(as.data.frame(props))                     %>% 
      add_colnames                                        %>% 
      {foo <- .; foo[1,] <- gsub('_', ' ', foo[1,]); foo} %>% 
      set_font(-1, everywhere, ss_font)                   %>% 
      set_bold(1, everywhere, TRUE)                       %>% 
      set_width(0.9)                                      %>% 
      set_background_color(everywhere, evens, grey(.9))   %>% 
      set_left_border(everywhere, 1, 1)                   %>% 
      set_right_border(everywhere, final(), 1)            %>% 
      set_top_border(1, everywhere, 1)                    %>% 
      set_bottom_border(1, everywhere, 1)                 %>% 
      set_bottom_border(final(), everywhere, 1)           %>% 
      set_top_padding(2)                                  %>% 
      set_bottom_padding(4)                               %>% 
      set_caption('Huxtable properties')                  %>% 
      set_position('left') %>% 
      set_col_width(c(.2, .25, .15, .15, .25))

prop_hux
```


## Tidyverse syntax

If you prefer a tidyverse style of code, using the pipe operator `%>%`, then you can use `set_*`
functions to set properties. These have the same name as the property, with `set_` prepended. For 
example, to set the `bold` property, you use the `set_bold` function.

`set_*` functions return the modified huxtable, so you can chain them together like this:

```{r}

library(dplyr)
hux(
        Employee     = c('John Smith', 'Jane Doe', 'David Hugh-Jones'), 
        Salary       = c(50000, 50000, 40000),
        add_colnames = TRUE
      )                               %>%
      set_right_padding(10)           %>%
      set_left_padding(10)            %>% 
      set_bold(1, 1:2, TRUE)          %>% 
      set_bottom_border(1, 1:2, 1)    %>%
      set_align(1:4, 2, 'right')      %>%
      set_number_format(2)            %>% 
      set_caption('Employee table')

```


`set_*` functions for cell properties are called like this: `set_xxx(ht, row, col, value)` or like
this: `set_xxx(ht, value)`. If you use the second form, then the value is set for all cells. `set_*`
functions for table properties are always called like `set_xxx(ht, value)`. We'll learn more about
this interface in a moment.

There are also some useful convenience functions:

* `set_all_borders` sets left, right, top and bottom borders for selected cells;
* `set_all_border_colors` sets left, right, top and bottom border colors;
* `set_all_border_styles` sets left, right, top and bottom border styles;
* `set_all_padding` sets left, right, top and bottom padding (the amount of space between the content
   and the border);
* `set_outer_borders` sets an outer border around a rectangle of cells.

## Getting properties

To get the current properties of a huxtable, just use the properties function without the left arrow:

```{r, results = 'markup'}
italic(ht)
position(ht)
```

As before, you can use subsetting to get particular rows or columns:

```{r, results = 'markup'}
bottom_border(ht)[1:2,]
```


# Editing content


## Changing text in a huxtable

You can treat a huxtable just like a data frame. If you want to change data in a cell, assign to
that cell:

```{r}
ht[3, 1] <- 'Jane Jones'
ht
```

To add a column, do, e.g.:

```{r}
ht_with_roles <- ht
ht_with_roles$Role <- c("Role", "Admin", "CEO", "Dogsbody")
ht_with_roles
```

There are two things to notice here:

* When we added the column, we included the column name explicitly in the data so it would be
  printed out.
* The third column doesn't have the properties we set on the first two columns, like the bold first
  row and the underlining.

If we want new columns to have the properties from their neighbours, we can use `cbind`, a base R
function that binds columns together. When you `cbind` huxtable objects, by default, cell properties
are copied over from their neighbours:

```{r}
ht_with_roles <- cbind(ht, c("Role", "Admin", "CEO", "Dogsbody"))
ht_with_roles
```

`rbind` works the same way:

```{r}
rbind(ht, c("Yihui Xie", 100000))
```

Notice how Yihui's salary has got the same number formatting as the other employees. That's because
cell properties for the new row were copied from the row above.

If you want to avoid this behaviour, use `copy_cell_props = FALSE`:

```{r}
rbind(ht, c("Yihui Xie", 100000), copy_cell_props = FALSE)
```


## Data manipulation the base R way

You can subset, sort and generally data-wrangle a huxtable just like a normal data frame. Cell and 
table properties will be carried over into subsets.

```{r}
# Select columns by name:
cars_mpg <- car_ht[, c("mpg", "cyl", "am")] 
# Order by number of cylinders:
cars_mpg <- cars_mpg[order(cars_mpg$cyl),]

cars_mpg <- cars_mpg                          %>% 
      huxtable::add_rownames(colname = "Car") %>% 
      huxtable::add_colnames()                %>%
      set_right_border(0.4)                   %>% 
      set_right_border_color("grey")

# Show the first 5 rows:
cars_mpg[1:5,]
```


## Data manipulation the dplyr way

You can also use `dplyr` functions to edit a huxtable:


```{r}
car_ht <- car_ht                                          %>%
      huxtable::add_rownames(colname = "Car")             %>%
      slice(1:10)                                         %>% 
      select(Car, mpg, cyl, hp)                           %>% 
      arrange(hp)                                         %>% 
      filter(cyl > 4)                                     %>% 
      rename(MPG = mpg, Cylinders = cyl, Horsepower = hp) %>% 
      mutate(kml = MPG/2.82)                               


car_ht <- car_ht                               %>% 
      set_number_format(1:7, "kml", 2)         %>% 
      set_col_width(c(.35, .15, .15, .15, .2)) %>% 
      set_width(.6)                            %>% 
      huxtable::add_colnames()                 %>% 
      set_right_border(0.4)                    %>% 
      set_right_border_color("grey")

car_ht
```

In general it is a good idea to prepare your data first, before styling it. For example, it was
easier to sort the `cars_mpg` data by cylinder, before adding column names to the data frame itself.

## Functions to insert rows, columns and footnotes

Huxtable has three convenience functions for adding a row or column to your table: `insert_row`,
`insert_column` and `add_footnote`.
`insert_row` and `insert_column` let you add a single row or column. The `after` parameter specifies
where in the table to do the insertion, i.e. after what row or column number.
`add_footnote` adds a single cell in a new row at the bottom. The cell spans the whole table row,
and has a border above.

```{r}
ht <- insert_row(ht, "Hadley Wickham", "100000", after = 3)
ht <- add_footnote(ht, "DHJ deserves a pay rise")
ht
```


# More formatting 

## Number format

You can change how huxtable formats numbers using `number_format`. Set `number_format` to a number
of decimal places (for more advanced options, see the help files). This doesn't just affect 
cells that are numbers: it works on any numbers within the cells. So, for example, if you have 
a cell like "12.00001 (3.0003)", `number_format` will affect both the numbers

```{r}
pointy_ht <- hux(c("Column heading", 11.003, 300, 12.02, "12.1 **", "mean 11.7 (se 2.3)")) %>% 
    set_all_borders(1)

number_format(pointy_ht) <- 3
pointy_ht
```

You can also align columns by decimal places. If you want to do this for a cell, just set the `align` property
to '.' (or whatever you use for a decimal point). 

```{r}
align(pointy_ht)[2:5, ] <- "." # not the first row
pointy_ht
```

There is currently no true way to align cells by the decimal point in HTML, and only limited
possibilities in TeX, so this works by right-padding cells with spaces. The output may look better
if you use a fixed width font.

## Automatic formatting

By default, when you create a huxtable using `huxtable` or `as_huxtable`, the package will guess
defaults for number formatting and alignment, based on the type of data in your columns. Numeric
data will be right-aligned or aligned on the decimal point; character data will be left aligned;
and the package will try to set sensible defaults for number formatting. If you want to, you can
turn this off with `autoformat = FALSE`:

```{r}

my_data <- data.frame(
        Employee           = c("John Smith", "Jane Doe", "David Hugh-Jones"), 
        Salary             = c(50000L, 50000L, 40000L),
        Performance_rating = c(8.9, 9.2, 7.8)  
      )
as_huxtable(my_data, add_colnames = TRUE) # with automatic formatting

as_huxtable(my_data, add_colnames = TRUE, autoformat = FALSE) # no automatic formatting
```

## Escaping HTML or LaTeX

By default, HTML or LaTeX code will be escaped:

```{r}
code_ht <- if (is_latex) hux(c("Some maths", "$a^b$")) else 
      hux(c("Copyright symbol", "&copy;"))
code_ht
```

To avoid this, set the `escape_contents` property to `FALSE`.

```{r}
escape_contents(code_ht)[2, 1] <- FALSE
code_ht
```


## Width and cell wrapping

You can set table widths using the `width` property, and column widths using the `col_width` property. If you use
numbers for these, they will be interpreted as proportions of the table width (or for `width`, a proportion of the
width of the surrounding text). If you use character vectors, they must be valid CSS or LaTeX widths. The only
unit both systems have in common is `pt` for points. 

```{r}
width(ht) <- 0.35
col_width(ht) <- c(.7, .3)
ht
```


It is best to set table width explicitly, then set column widths as proportions.

By default, if a cell contains long contents, it will be stretched. Use the `wrap` property to allow cell contents
to wrap over multiple lines:

```{r}
ht_wrapped <- ht
ht_wrapped[5, 1] <- "David Arthur Shrimpton Hugh-Jones"
wrap(ht_wrapped) <- TRUE
ht_wrapped
```


## Adding row and column names

Just like data frames, huxtables can have row and column names. Often, we want to add these to the final table.
You can do this using either the `add_colnames`/`add_rownames` arguments to `as_huxtable`, or the
`add_colnames()`/`add_rownames()` functions. (Note that earlier versions of `dplyr` used to have functions with the
same name.)

```{r}
as_hux(mtcars[1:4, 1:4])                           %>% 
      huxtable::add_rownames(colname = "Car name") %>% 
      huxtable::add_colnames()
```


## Merging cells

Sometimes you want a single cell to spread over more than one row or column: for example, if you
want a heading that covers several different rows.

You can do this by calling `merge_cells(ht, rows, cols)`. `rows` and `cols` should be a contiguous
sequence of numbers. The rectangle of cells `ht[rows, cols]` will be merged. 

When cells in a rectangle are merged, all cells apart from the top left one are hidden, along with
any properties they have. So if you want to set cell properties, you have to target the top left 
cell.

Here, we'll add some row and column headings to the `mtcars` dataset:

```{r}
cars_mpg <- cbind(cylinders = cars_mpg$cyl, cars_mpg)
cars_mpg$cylinders[1]   <- ""
cars_mpg$cylinders[2]   <- "Four cylinders"
cars_mpg$cylinders[13]  <- "Six cylinders"
cars_mpg$cylinders[20]  <- "Eight cylinders"

cars_mpg <- cars_mpg %>%  
  merge_cells(2:12, 1) %>% 
  merge_cells(13:19, 1) %>% 
  merge_cells(20:33, 1)

cars_mpg <- rbind(c("List of cars", "", "", "", ""), cars_mpg)
cars_mpg <- merge_cells(cars_mpg, 1, 1:5)
align(cars_mpg)[1, 1] <- "center"

# a little more formatting:

cars_mpg <- set_all_padding(cars_mpg, 2)
cars_mpg <- set_all_borders(cars_mpg, 1)
valign(cars_mpg)[1,] <- "top"
col_width(cars_mpg) <- c(.4 , .3 , .1, .1, .1)
number_format(cars_mpg)[, 4:5] <- 0
bold(cars_mpg)[1:2, ] <- TRUE
bold(cars_mpg)[, 1] <- TRUE
if (is_latex) font_size(cars_mpg) <- 10
cars_mpg
```

`merge_cells` works by setting the top left cell's `colspan` and `rowspan` properties. If you know HTML tables, then 
these will be familiar to you. `colspan` sets how many columns the cell covers, and `rowspan` sets
how many rows the cell covers. If you prefer, you can set these directly:

```{r, eval = FALSE}
colspan(cars_mpg)[1, 1] <- 5
```


## Quick themes

Huxtable comes with some predefined themes for formatting.

```{r}
theme_plain(car_ht)
```


# Selecting rows, columns and cells


## Row and column functions

If you use the `set_*` style functions, huxtable has some convenience functions for selecting rows and columns.

To select all rows, or all columns, use `everywhere` in the row or column specification. To select just even or odd-numbered rows or columns, use `evens` or `odds`. To select the last `n` rows or columns, use `final(n)`.
To select every *n*th row, use `every(n)` and to do this starting from row *m* use `every(n, from = m)`.

With these functions it is easy to add striped backgrounds to tables:

```{r}
car_ht                                                 %>% 
      set_background_color(evens, everywhere, "wheat") %>% 
      set_background_color(odds, everywhere, grey(.9)) %>% 
      set_bold(1, everywhere, TRUE)
```


Of course you could also just do `1:nrow(car_ht)`, but, in the middle of a dplyr pipe, you may not know exactly
how many rows or columns you have. Also, these functions make your code easy to read.

You can also use `dplyr` functions like `starts_with()`, `contains()`, and `matches()` to specify columns by column 
name. For a full list of these functions, see `?select_helpers`.

```{r}
car_ht %>% set_background_color(everywhere, starts_with("C"), "orange")
car_ht %>% set_italic(everywhere, matches("[aeiou]"), TRUE)
```

Note that unlike in `dplyr`'s `select` function, you have to specify rows as well as columns.

Lastly, remember that you can set a property for every cell by simply omitting the `row` and `col` arguments, like this:
`set_background_color(ht, 'orange')`.


## Conditional formatting

You may want to apply conditional formatting to cells, based on their contents. Suppose we want to display a table of correlations, and to highlight ones which are significant. We can use the `where()` function to select those cells.

```{r}
library(psych)
data(attitude)
att_corr <- corr.test(as.matrix(attitude))

att_hux <- as_hux(att_corr$r)                                           %>% 
      # selects cells with p < 0.05:
      set_background_color(where(att_corr$p < 0.05), "yellow")          %>% 
      # selects cells with p < 0.01:
      set_background_color(where(att_corr$p < 0.01), "orange")          %>% 
      set_text_color(where(row(att_corr$r) == col(att_corr$r)), "grey") 


att_hux <- att_hux                                                      %>% 
      huxtable::add_rownames()                                          %>% 
      huxtable::add_colnames()                                          %>%
      set_caption('Correlations in attitudes among 30 departments')     %>% 
      set_bold(1, everywhere, TRUE)                                     %>% 
      set_bold(everywhere, 1, TRUE)                                     %>% 
      set_all_borders(1)                                                %>%
      set_number_format(2)                                              %>% 
      set_position('left')

att_hux

```

We have now seen three ways to call `set_*` functions in huxtable:

* With four arguments, like `set_property(hux_object, rows, cols, value)`;
* With two arguments, like `set_property(hux_object, value)` to set a property everywhere;
* With three arguments, like `set_property(hux_object, where(condition), value)` to set a property for specific cells.

The second argument of the three-argument version must return a 2-column matrix. Each row of the matrix gives one cell.
`where()` does this for you: it takes a logical matrix argument and returns the rows and columns where a condition is
`TRUE`. It's easiest to show this with an example:

```{r}
m <- matrix(c('dog', 'cat', 'dog', 'dog', 'cat', 'cat', 'cat', 'dog'), 4, 2)
m
where(m == 'dog') # m is equal to 'dog' in cells (1, 1), (3, 1), (4, 1) and (4, 2):
```

`set_*` functions have one more optional argument, the `byrow` argument, which is `FALSE` by default. If you set a
single pattern for many cells, you may want the pattern to fill the matrix by column or by row. The default fills the
pattern in going down columns. If you set `byrow = TRUE`, the pattern goes across rows instead. (This is a bit
confusing: typically, `byrow = TRUE` means that the *columns* will all look the same. But it works the same way as the
`byrow` argument to `matrix()`.)

```{r}

color_demo <- matrix('text', 7, 7)
rainbow <- c('red', 'orange', 'yellow', 'green', 'blue', 'turquoise', 'violet')
color_demo <- as_hux(color_demo)                  %>% 
      set_text_color(rainbow)                     %>% # text rainbow down columns
      set_background_color(rainbow, byrow = TRUE) %>% # background color rainbow along rows
      set_all_borders(1)                          %>% 
      set_all_border_colors('white')
color_demo

```



# Creating a regression table

A common task for scientists is to create a table of regressions. The function `huxreg` does this for you. Here's a quick example:

```{r}
data(diamonds, package = 'ggplot2')

lm1 <- lm(price ~ carat, diamonds)
lm2 <- lm(price ~ depth, diamonds)
lm3 <- lm(price ~ carat + depth, diamonds)

huxreg(lm1, lm2, lm3)
```

For more information see the `huxreg` vignette, available online in
[HTML](https://hughjonesd.github.io/huxtable/huxreg.html) or [PDF](https://hughjonesd.github.io/huxtable/huxreg.pdf) or
in R via `vignette('huxreg')`.

# Output to different formats

## Automatic pretty-printing of data frames

If you load huxtable within a knitr document, it will automatically format data frames for you by
installing a `knit_print.data_frame` command.


```{r, include = FALSE}
options(huxtable.knit_print_df = TRUE)
```

```{r}
head(mtcars)
```

If you don't want this (e.g. if you want to use `knitr::kable` or the [printr package](https://cran.r-project.org/package=printr/vignettes/printr.html), then you can
turn it off like this:

```{r}
options(huxtable.knit_print_df = FALSE)

head(mtcars) # back to normal

options(huxtable.knit_print_df = TRUE)
```

## Using huxtables in knitr and rmarkdown

If you use knitr and rmarkdown in RStudio, huxtable objects should automatically display in the
appropriate format (HTML, LaTeX or RTF). You need to have some LaTeX packages installed for huxtable
to work in LaTeX. To find out what these are, you can call `report_latex_dependencies()`. This will
print out and/or return a set of `usepackage{...}` statements. If you use Sweave or knitr without
rmarkdown, you can use this function in your LaTeX preamble to load the packages you need. You can
also automatically install these packages using `install_latex_dependencies()`.

Rmarkdown exports to Word via Markdown. You can use huxtable to do this, but since Markdown tables
are rather basic, a lot of formatting will be lost. If you want to create Word or Powerpoint
documents directly, install the [flextable package](https://cran.r-project.org/package=flextable)
from CRAN. You can then convert your huxtable objects to `flextable` objects and include them in
Word or Powerpoint documents. Almost all formatting should work. See the `flextable` and `officer`
documentation and `?as_flextable` for more details.

Similarly, to create formatted reports in Excel, install the [openxlsx
package](https://cran.r-project.org/package=openxlsx). You can then use `as_Workbook` to convert
your huxtables to Workbook objects, and save them using `openxlsx::saveWorkbook`.

Sometimes you may want to select how huxtable objects are printed by default. For example, in an
RStudio notebook (a .Rmd document with `output_format = html_notebook`), huxtable can't
automatically work out what format to use, as of the time of writing. You can set it manually using
`options(huxtable.print = print_notebook)` which prints out HTML in an appropriate format.

You can print a huxtable on screen using `print_screen` (or just by typing its name at the command
line.) Borders, column and row spans and cell alignment are shown. If the
[crayon](https://cran.r-project.org/package=crayon) package is installed, and your terminal or R IDE
supports it, border, text and background colours are also displayed.

```{r, results = 'markup'}
print_screen(ht)
```


If you need to output to another format, file an 
[issue request](https://github.com/hughjonesd/huxtable) on Github. 

## Quick output commands

Sometimes you quickly want to get your data into a document. To
do this you can use huxtable functions starting with `quick_`, shown below.

```{r, echo = FALSE}
quick_commands <- hux(
        Command = paste0("<code>", 
          c("quick_pdf", "quick_docx", "quick_html", "quick_xlsx", "quick_pptx", "quick_rtf"), 
          "</code>"),
        Output = c("PDF document", "Word document", "HTML web page", "Excel spreadsheet", 
          "Powerpoint presentation", "RTF document"),
        add_colnames = TRUE
      )
escape_contents(quick_commands) <- FALSE

theme_plain(quick_commands)
```

These are called with one or more huxtable objects (or objects which can be turned into a huxtable,
such as data frames). A new document of the appropriate type will be created and opened. By default
the file will be in the current directory, under a name like e.g. `huxtable-output.pdf`. If the file
already exists, you'll be asked to confirm the overwrite. For non-interactive use, you must specify
a filename yourself explicitly. This keeps you from accidentally trashing your files.

```{r, eval = FALSE}
quick_pdf(mtcars) 
quick_pdf(mtcars, file = 'motorcars data.pdf')
```

# End matter

For more information, see the [website](https://hughjonesd.github.io/huxtable) or 
[github](https://github.com/hughjonesd/huxtable).
