---
title: "Introduction to Huxtable"
author: "David Hugh-Jones"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to Huxtable}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

Huxtable is a package for writing LaTeX and HTML tables. It is powerful, but easy to use. It is meant to be a
replacement for packages like xtable, which are useful but not always very user-friendly.

To create a table with huxtable, use the function `huxtable`, or `hux` for short.

```{r}
library(huxtable)
ht <- hux(Employee = c('John Smith', 'Jane Doe', 'David Hugh-Jones'), Salary = c(50000, 50000, 40000),
      add_colnames = TRUE)
```


Or, if you already have your data in a data frame, use `as_hux`.

```{r}
data(mtcars)
car_ht <- as_hux(mtcars, add_colnames = TRUE)
```

Huxtables are simply data frames, along with some extra information on how to display them. If you look at them in R,
they'll appear just like ordinary data frames.

```{r}
print(ht)
```

If you evaluate them in a knitr document (like this one!) they print as tables, using HTML or LaTeX as appropriate.

```{r}
ht
```

The default output is a very plain table. To customize it, you can set various properties. Let's make our table headings bold, draw a line under the header row, and align the second row to the right:


```{r}
bold(ht)[1,] <- TRUE
bottom_border(ht)[1,] <- TRUE
align(ht)[,2] <- 'right'
```

Vignettes are long form documentation commonly included in packages. Because they are part of the distribution of the package, they need to be as compact as possible. The `html_vignette` output type provides a custom style sheet (and tweaks some options) to ensure that the resulting html is as small as possible. The `html_vignette` format:

- Never uses retina figures
- Has a smaller default figure size
- Uses a custom CSS stylesheet instead of the default Twitter Bootstrap style

## Vignette Info

Note the various macros within the `vignette` section of the metadata block above. These are required in order to instruct R how to build the vignette. Note that you should change the `title` field and the `\VignetteIndexEntry` to match the title of your vignette.

## Styles

The `html_vignette` template includes a basic CSS theme. To override this theme you can specify your own CSS in the document metadata as follows:

    output: 
      rmarkdown::html_vignette:
        css: mystyles.css

## Figures

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))
