---
title: "Untitled"
format: docx
---

## Table for word document in quarto

```{r}
library(huxtable)
library(officedown)
```

```{r}
#| label: tbl-air1

mytab1 <- hux(head(airquality))
caption(mytab1)<- "My caption"
mytab1
```

- Table caption appears
- This reference @tbl-air1 doesn't work

```{r}
mytab2 <- hux(head(airquality))
caption(mytab2) <- "Second table caption"
label(mytab2) <- "tbl-air2"
mytab2
```

- Table caption appears
- This reference @tbl-air2 doesn't work

```{r}
#| label: tbl-air3
#| tbl-cap: "Third table caption"

mytab3 <- hux(head(airquality))
mytab3
```

- Table caption doesn't appear
- This reference @tbl-air3 doesn't work



Testing flextable on its own: `r officer::run_reference(id = "tbl-flextable")`.

```{r}
#| label: tbl-flextable
#| tbl-cap: Caption **strong**
flextable::flextable(mtcars[1:4, 1:4])
```


```{r}
#| label: tbl-air4
#| tbl-cap: "Fourth table caption"

mytab4 <- hux(head(airquality))
label(mytab4) <- "tbl-air4"
mytab4
```

- Table caption?
- Using run_reference workaround, does reference work? `r officer::run_reference(id = "tbl-air4")` 
