---
title: "Example of cross-references and captions"
output: 
  bookdown::html_document2
---

# The first chapter {#firstchap}

This is the first chapter with id `{#sec:firstchap}`

## A subchapter {#sec:subchapter}

* Anchors have curly braces. 
* This is a subchapter with id `{#sec:subchapter}`.
* You can also link to [arbitrary text]{#atext}. Anchor looks like this: `[arbitrary text]{#atext}`. The hash is required.

## Another one

This chapter only exists to demonstrate implicit anchors. You should prefer the above explicit ones.

```{r, echo = FALSE, warning=FALSE}
library(knitr)
library(pander)
library(DT)
library(kableExtra)
library(ggplot2)
mcaption = function(tag, caption){
  # Do not use underscores in tag!
  stopifnot(length(grep("_", tag)) == 0)
  cat("<caption>(#tab:", tag, ")", caption, "</caption>", sep="")
}
```

### `kable` Table

Too bad, caption is always at the top

```{r iris}
# Chunk iris, not tab:iris!. Gives Table 1.1
kable(head(iris), caption = "Iris with kable")
kable(head(iris), caption = "Iris with kable")
```

### `kableExtra` Table 

`kableExtra` also plays well with numbering and captions.

```{r irisextra}
kable(head(iris), caption = "Iris with kableExtra") %>% 
   kable_styling()
```

### `pander` Table
```{r panderiris, results='asis', echo = FALSE}
# Must use echo = FALSE (or globally). 
mcaption("pandertable", "This is a pander caption")
pander(head(iris))
```

### `DT` Table
```{r datatableiris, results='asis', echo=FALSE}
mcaption("dtiris", "This is a DT")
DT::datatable(head(iris), caption = "Iris with pander")
```

### `flextable`

Currently requires a version from github (https://github.com/davidgohel/flextable/pull/205). This was tested with `flextable` version `r packageVersion("flextable")`.

```{r}
use_flex = TRUE # set to FALSE to test huxtable
```


```{r flexiris, echo=FALSE, eval = use_flex}
suppressPackageStartupMessages(library(flextable))
ft =  flextable( head( iris ) )
ft = flextable::set_caption(ft, "Iris with flextable")
ft
```

### `huxtable`

`huxtable` extends `flextable`, but cannot be used together with `flextable` because of a naming conflict. When used alone, it does not show autonumbering. This was tested with `huxtable` version `r packageVersion("huxtable")`.

```{r huxiris, echo=FALSE}
suppressPackageStartupMessages(library(huxtable))
hx =  hux( head( iris ) ) 
hx = add_colnames(hx)
hx = huxtable::set_caption(hx, "Iris with huxtable")
hx = huxtable::set_label(hx, "tab:huxiris")
hx
```

### A figure 
```{r irisf, fig.height=2.4, fig.width= 2.5, fig.cap="Figures work nicely", echo = FALSE}
ggplot(iris, aes(x=Sepal.Length, y=Sepal.Width) ) + geom_point()
```

\begin{equation}
\bar{X} = \frac{\sum_{i=1}^n X_i}{n} (\#eq:mean)
\end{equation}

## References to tables, figures and equations

Currently, referencing tables and autonumbering only works for `kable`, not for pander or DT, promised for `flextable` and its descendant `huxtable`. 

* Explicit reference to chapter: Check also Chapter \@ref(firstchap)...
* Explicit reference to chapter with optional category: Check also Chapter \@ref(sec:subchapter).
* Implicit reference: As Chapter \@ref(another-one) shows... can only be used when there is no explicit anchor.
* Table: `kable` table \@ref(tab:iris) is best seen here. The `tab:` is implicit, the chunk must not have the `tab:` prefix.
* Reference to a `kableExtra`table:  As the extra Table \@ref(tab:irisextra) shows ....
* Reference to `flextable`: Please look at flextable \@ref(tab:flexiris).
* Fig. \@ref(fig:irisf) shows that everything works nicely with figures. Great work by Yihui.
* Equation: Also see Equation \@ref(eq:mean).

## Links to tables and figures

* [An internal link](#atext) to arbitrary text. `[An internal link](#atext)`. The hash is required.
* [An internal link](#sec:subchapter) to a subchapter.
* [Link to figure](#fig:irisf), a link to a figure `[Link to figure](#fig:irisf)`, the hash is required.
* [Link to flextable](#tab:flexiris)
* [Link to huxtable](#tab:huxiris); does not work currently.

For `pander`  a link to the chunk label fails `\@ref(panderiris)`, but you can use the `mcaption` tag using the function defined in this document. Do not forget to put `results='asis'` into the chunk header.

* You should check Table \@ref(tab:pandertable).... Don't use underscores in tag!
* [Link to table](#tab:pandertable). Must use hash here, because it is part of the `mcaption` function.
