---
title: "Example"
output: 
  pdf_document:
    keep_tex: true
---

```{r setup, include=FALSE}
library(magrittr)
library(huxtable)
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

lorem upsdf usf ueriawo fsd urweia dsjkl ewauof dsnkfea uieorw lorem upsdf usf
ueriawo fsd urweia dsjkl ewauof dsnkfea uieorw lorem upsdf usf ueriawo fsd
urweia dsjkl ewauof dsnkfea uieorw lorem upsdf usf ueriawo fsd urweia dsjkl
ewauof dsnkfea uieorw lorem upsdf usf ueriawo fsd urweia dsjkl ewauof dsnkfea
uieorw lorem upsdf usf ueriawo fsd urweia dsjkl ewauof dsnkfea uieorw


```{r}

ht <- huxtable::huxtable(a = "L L L L L L L left L L L L L L L", 
                         b = "R R R R R R R rght R R R R R R")

huxtable::position(ht) <- "center"

huxtable::width(ht) <- 0.5 # This can be varied
ht <- huxtable::set_all_borders(ht, 0.4)

huxtable::wrap(ht) <- TRUE
ht
```


\color{red}\rule{0.1\linewidth}{0.5pt}\color{blue}\rule{0.1\linewidth}{0.5pt}\color{red}\rule{0.1\linewidth}{0.5pt}\color{blue}\rule{0.1\linewidth}{0.5pt}\color{red}\rule{0.1\linewidth}{0.5pt}\color{blue}\rule{0.1\linewidth}{0.5pt}\color{red}\rule{0.1\linewidth}{0.5pt}\color{blue}\rule{0.1\linewidth}{0.5pt}\color{red}\rule{0.1\linewidth}{0.5pt}\color{blue}\rule{0.1\linewidth}{0.5pt}

```{r}
ht %>% set_all_padding(8)
```
