---
title: "Huxtable functionality demo/test"
author: "David Hugh-Jones"
date: "2/03/2017"
output:
  html_document: null
  pdf_document:
    includes:
      in_header: placeins-header.tex
    keep_tex: yes
---


```{r setup, echo = FALSE}
knitr::asis_output('', meta = rmarkdown::latex_dependency('placeins'))
library(huxtable)
is_latex <- guess_knitr_output_format() == 'latex'

knitr::knit_hooks$set(
  barrier = function(before, options, envir) {
    if (! before && is_latex) knitr::asis_output('\\FloatBarrier')
  }
)
knitr::opts_chunk$set(echo = TRUE, barrier = TRUE)

```

# Basic usage


## Creating a huxtable

```{r}
ht_orig <- huxtable(a = c('Parsley', 'Sage', 'Rosemary', 'Thyme'), b = 10 ^ (2:5)) 
ht_orig <- set_all_borders(ht_orig, 1:nrow(ht_orig), 1:ncol(ht_orig), 1)
ht_orig
```


## Borders

```{r}
ht <- ht_orig
ht <- set_all_borders(ht, 1:4, 1:2, 0)
top_border(ht) <- c(2, 0, 0, 1)
bottom_border(ht)[4,] <- 2 
ht

```

## Cell alignment

```{r}
ht <- ht_orig
align(ht)[,1] <- 'left'
ht

```

## Vertical alignment

```{r}
ht <- ht_orig
row_height(ht) <- '50pt'
valign(ht)[,1] <- 'top'
valign(ht)[,2] <- 'bottom'
ht

```

## Column span

```{r}

ht <- ht_orig
background_color(ht)[1,1] <- 'green'
colspan(ht)[1,1] <- 2
ht   

```

## Row span

```{r}
ht <- ht_orig
ht[1,1] <- 'A rather long block of text'
background_color(ht)[1,1] <- 'green'
rowspan(ht)[1,1] <- 2
ht   

```

## Table position

```{r}
ht <- ht_orig
position(ht) <- 'left'
ht
```

## Table width

```{r}
ht <- ht_orig
width(ht) <- 1/3
ht

```

## Column width

```{r}
ht <- ht_orig
col_width(ht) <- c(.8, .2)
ht

```

## Row height

```{r}
ht <- ht_orig
height(ht) <- 0.2; 
if (! is_latex) height(ht) <- '100px' # need a specific height for row heights to work in HTML
row_height(ht) <- c(.4, .2, .2, .2) * 0.2
ht

```

## Text color, bold, italic, font size

```{r}
ht <- ht_orig
text_color(ht)[,2] <- 'red'
bold(ht)[1,] <- TRUE
italic(ht)[3,] <- TRUE
font_size(ht)[2,] <- 14
font(ht)[4,] <- if (is_latex) 'phv' else 'times' 
ht

```

## Number formatting

```{r}
ht <- ht_orig
ht[,2] <- ht[,2] + rnorm(4)
number_format(ht)[2,] <- 2
number_format(ht)[3,] <- '%011.4f'
number_format(ht)[4,] <- list(function(x) prettyNum(round(x, 3), big.mark = ','))

ht

```

## Replace NAs

```{r}
ht <- ht_orig
ht[2,] <- NA
ht
na_string(ht) <- '--'
ht

```

## Pipe style with magrittr

```{r}

if (require('magrittr')) {
ht <- ht_orig
ht %>% 
      set_all_borders(1:4, 1:2, 0)            %>% 
      set_top_border(1, 1:2, 1)               %>% 
      set_bold(1, 1:2, TRUE)                  %>% 
      set_background_color(1:4, 1:2, 'wheat') %>% 
      set_number_format(1:4, 1:2, '%03.1d')   %>% 
      set_align(1:4, 1, 'left')               %>% 
      set_align(1:4, 2, 'right')
}

```

## Quick themes

```{r}
ht <- ht_orig
ht <- rbind(c("Herb", "Amount"), ht)
theme_minimal(ht)
```

```{r}
theme_striped(ht, header_col = FALSE)
```

```{r}
theme_article(ht)
```


# Converting other objects to huxtables

```{r}
defaults <- function(ht) {
  bottom_border(ht)[1,] <- 1
  background_color(ht)[seq(1, nrow(ht), 2), ] <- grey(.9)
  ht
}
```

```{r}
defaults(as_hux(letters[1:5]))
defaults(as_hux(matrix(letters[1:10], 2)))
```

```{r}
data(mtcars)
defaults(as_hux(mtcars[1:10,]))
```

```{r}
car_table <- xtabs(~ cyl + gear, mtcars)
defaults(as_hux(car_table))
```

```{r}
car_ft <- ftable(cyl ~ gear + vs, mtcars)
defaults(as_hux(car_ft))
```


# Joining, subsetting and manipulating huxtables

## Subsets

```{r}
ht <- ht_orig
bottom_border(ht)[c(1,4),] <- 1
background_color(ht)[1,] <- 'wheat'

ht[1:3,1]
```

## Joining

```{r}

ht2 <- ht_orig
italic(ht2) <- TRUE
rbind(ht, ht2)
cbind(ht, ht2)
rbind(c("Oregano", 300), ht)
```

## Transpose

```{r}
colspan(ht)[1, 1] <- 2
ht
t(ht)
```


# Advanced features

## Printing on screen

```{r}
ht <- ht_orig
# Basic data:
print(ht)

# Pretty print:
print_screen(ht)

# Markdown:
print_md(ht)
```

## Padding

```{r}
ht <- ht_orig
ht <- set_all_padding(ht, 1:4, 1:2, 0)
left_padding(ht)[,1] <- 40
right_padding(ht)[,2] <- 40
top_padding(ht) <- 20
bottom_padding(ht) <- 20
ht

```

## Cell rotation

```{r}
ht <- ht_orig
height(ht) <- 0.2 # necessary
if (! is_latex) height(ht) <- '300px'
col_width(ht) <- c(.25, .75)
rotation(ht)[,1] <- 90
ht

```

## Border width

```{r}
ht <- ht_orig
bottom_border(ht)[1,] <- 3
ht

```


## Caption

```{r}
ht <- ht_orig
caption(ht) <- 'A simple table'
ht

```


## Caption below

```{r}
ht <- ht_orig
caption(ht) <- 'A simple table'
caption_pos(ht) <- 'bottom'
ht

```


## Label

```{r}
ht <- ht_orig
caption(ht) <- 'Captions are required for labels to work'
label(ht) <- 'tab:mytable'
ht
knitr::asis_output('If this is LaTeX we can see a reference to table \\ref{tab:mytable}.')
```


NB: references may not work if knitting PDFs in Rstudio. 


# Testing corner cases

## Heights and padding

```{r}
ht <- ht_orig

row_height(ht) <- c(.1, .05, .05, .05)
# top_padding(ht) <- 10
# bottom_padding(ht) <- 20
ht
```

## Row and column span

```{r}
ht <- ht_orig
ht[1,1] <- 'A rather long block of text'
background_color(ht)[1,1] <- 'green'
rowspan(ht)[1,1] <- 2
colspan(ht)[1,1] <- 2
ht   

```

# Huge table of cross-validation

```{r}
set_rowcol <- function(ht, fun, rc, value) {
  ht[1, rc] <- ht[rc, 1] <- paste(deparse(substitute(fun)), '=', value)
  ht <- fun(ht, rc, 1:ncol(ht), value)
  ht <- fun(ht, 1:nrow(ht), rc, value)
  ht
}

N <- 23
megahux <- as_hux(matrix("", N, N))
megahux[1:5,] <- 'Some text'
megahux[6:10,] <- rnorm(5*N)
megahux[11:15,] <- 'Some more text'
megahux[18:23, ] <- rnorm(6 * N)
megahux[16:17, ] <- substring(stringi::stri_rand_lipsum(2*N), 0, 50)
# for (col in c(2,7,12,17,22)) test_hux[,col] <- sample(test_hux[,col])


megahux <- megahux                                %>% 
      set_rowcol(set_valign, 1, 'top')              %>% 
      set_rowcol(set_align, 2, 'right')              %>% 
      set_rowcol(set_top_border, 3, 1)              %>% 
      set_rowcol(set_bottom_border, 4, 2)           %>% 
      set_rowcol(set_left_border, 5, 1)             %>% 
      set_rowcol(set_right_border, 6, 1)            %>% 
      set_rowcol(set_background_color, 10, 'wheat')  %>% 
      set_rowcol(set_text_color, 11, 'red')         %>% 
      set_rowcol(set_top_padding, 12, 20)           %>% 
      set_rowcol(set_bottom_padding, 13, 20)        %>% 
      set_rowcol(set_left_padding, 14, 20)          %>% 
      set_rowcol(set_right_padding, 15, 20)         %>% 
      set_rowcol(set_escape_contents, 16, FALSE)    %>%
      set_rowcol(set_na_string, 17, 'N.a.')         %>%
      set_rowcol(set_bold, 18, TRUE)                %>%
      set_rowcol(set_italic, 19, TRUE)              %>%
      set_rowcol(set_font_size, 20, 20)             %>%
      set_rowcol(set_number_format, 21, '%9.6f')    %>%
      set_rowcol(set_font, 22, 'times')             %>% 
      set_rowcol(set_wrap, 23, TRUE)

colspan(megahux)[4, 5] <- 6
rowspan(megahux)[5, 4] <- 6
megahux[5,4] <- stringi::stri_rand_strings(1, 60)
megahux[4, 5] <- stringi::stri_rand_strings(1, 60)
wrap(megahux)[4, 5] <- wrap(megahux)[5, 4] <- TRUE
megahux[,16] <- megahux[16,] <- '<b>Non escaped HTML</b>'
megahux[,17] <- megahux[17,] <- sample(c(1:10, rep(NA, 13)))

mh <- list()
j <- 1
for (row in list(2:11, 12:23)) for (col in list(2:3, 4:10, 11:16, 17:23)) {
  mh[[j]] <- knitr::knit_print(megahux[c(1, row), c(1, col)])
  j <- j + 1
}

mh[[1]]
mh[[2]]
mh[[3]]
mh[[4]]
mh[[5]]
mh[[6]]
mh[[7]]
mh[[8]]


```


