---
title: "Reporting data with huxtable 4.3.0"
output: rmdformats::material
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Dealing with masses of data

Sometimes there is a lot of data to present. For example, you might have a big
table of regressions, or sales figures from several business units. You want to show 
the data, while highlighting certain key elements.

As an example, we'll use the `mtcars` dataset of 32 1970s automobiles.

```{r}
head(mtcars)
```


I will use the [huxtable
library](https://hughjonesd.github.io/huxtable) to create a formatted table, and the
dplyr pipe command to simplify transforming data.

```{r, message = FALSE}
library(dplyr)
library(huxtable)

car_hux <- as_huxtable(mtcars, add_colnames = TRUE, add_rownames = "Model")

car_hux <- set_font_size(car_hux, 9)

car_hux
```


This dataset contains many numbers and columns. It's easy to get lost. 
To start, let's give alternate rows a different background. That will help readers to
navigate through the data.

Huxtable 4.3.0 introduces `map_` commands. These let you map different data cells to
different display properties. Let's use `map_background_color` to set the background
colour.

```{r}

top <- function(data, n = 8) data %>% 
      head(n) %>% 
      insert_row(rep("...", ncol(.)), after = n)

car_hux %>% 
      top() %>% 
      map_background_color(by_rows("grey90", "grey98"))
```

Huxtable's built-in `theme_grey` will do that for us, and also set a few other
sensible cosmetic defaults. We'll use that as a basis from now on.

```{r}
car_hux <- theme_grey(car_hux, header_col = FALSE)
car_hux %>% top()
```


# Mapping data with `map_` commands

`map_` commands take a huxtable, an optional row and column specifier, and a "mapping function" which states how to map table data to output. Mapping functions start with
`by_`. For example, `by_rows` just says: apply this output to rows in sequence.

Let's use another mapping function to highlight the cars with the best and worst fuel efficiency.

```{r}

car_hux %>% map_text_color(everywhere, "mpg", 
      by_quantiles(c(0.2, 0.8), c("red", "black", "green3")))
```


`by_quantiles` takes a vector of quantiles, and a vector of colours. Here, we've coloured
the bottom 20% of data red, and the top 20% green. Notice that we used one more
colour name than quantile.

This example also showed how we can select rows or columns to map, using
the row and column specifier arguments to a `map_` command. The row specifier
was `everywhere`, which is just huxtable shorthand for `1:nrow(car_hux)`. The
column specifier was the column name.

# Map numbers in a range with `by_colorspace`

To map a continuous range of numbers, use `by_colorspace`. This takes
a vector of colours, then maps the data continuously between them. Here,
we'll highlight the "qsec" column, which reports the car's time to reach a quarter of a
mile. Lower is better (if you're an adrenalin junkie), so let's make low numbers
pop out with a yellow background:

```{r}

car_hux %>% map_background_color(everywhere, "qsec",
      by_colorspace("yellow", "red"))

```

# Highlight text with `by_regex`

`map_text_color` is another `map_` command. There is one of these commands for every
huxtable cell property. Another one is `map_bold`. Let's use this to pick out
the Mercedes models.

For character data, we can use `by_regex` to match certain cells: 

```{r}

car_hux %>% 
      map_bold(by_regex("Merc" = TRUE)) %>% 
      top(15)
```

The syntax of `by_regex` is `"match1" = value2, "match2" = value2, ...`. Each "match"
is a regular expression. You can read more about them in `?regexpr`. Here, we just
used a simple regular expression, which matched every cell containing the string 
`"Merc"`. 

# Highlight a column with `set_background_color` and `set_outer_borders`

Sometimes you might want to highlight an entire column of data, for example as 
you flip through a report. Here's one way to do this:

```{r}

car_hux %>% 
      set_background_color(everywhere, "hp", "yellow") %>% 
      top()
```

While `map_` commands map different cells to different values, `set_` commands
set all the cells in a range to the same value. Here again, we've used a 
row specifier `everywhere` to select all rows, and picked out the horsepower
column "hp" by name.

We can do the same trick to highlight rows. Here, we'll highlight the "Hornet" cars:

```{r}
car_hux %>% 
      set_background_color(grepl("Hornet", .$Model), everywhere, "yellow") %>% 
      top()
```

Another way to pick out a column is to set a border around it. We can do this
using the `set_outer_borders` function.

```{r}

car_hux %>% 
      set_all_borders(everywhere, "disp", 0) %>% 
      set_outer_borders(everywhere, "disp", 2) %>% 
      set_all_border_colors(everywhere, "disp", "darkred") %>%  
      top()
```

`set_outer_borders` sets borders around a square of cells. One wrinkle is that there
is not (yet) a `set_outer_border_colors` function. Instead, we had to set borders to
0 in the `"disp"` column, then set all border colours to dark red.

# More ways to map

To recap, here are some ways to highlight data using huxtable:

* Use `by_rows` to set properties on alternating rows.
* Use `by_quantiles` to highlight top and bottom numbers.
* `by_regex` picks out cells matching certain text.
* `set_background_color` and `set_outer_borders` can highlight whole rows and columns.

There are several other ways to map data. For example, `by_cols` sets cell properties by columns. `by_cases` and `by_function` are more general solutions to pick
out cells using more complex criteria. 

Huxtable 4.3.0 is available on [CRAN](https://cran.r-project.org/package=huxtable). Happy mapping!
  
